---
title: "Regression for Larralde & Noveck E2 replication"
author: Morgan Moyer
date: 7 Nov, 2022
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(lme4)
library(lmerTest)
library(multcomp) # not available for this version of R
library(philentropy)
library(stringr)
library(textstem)
library(tidyverse)
theme_set(theme_bw())
cbPalette <- c("#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73")
```

```{r setwd, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(this.dir)
```


# First looking at just single-processed data (no overal trial RT outlier removal)

```{r read in all the data, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
d.rt1 <- read.csv("../data/rt1_processed.csv")
d.rt2 <- read.csv("../data/rt2_processed.csv")
```


```{r}

# Make everything into a factor
d.rt1$Connective = as.factor(d.rt1$Connective)
d.rt1$ID = as.factor(d.rt1$ID)
d.rt1$TrialID = as.factor(d.rt1$TrialID)
d.rt1$ReactionTime1 = as.numeric(d.rt1$ReactionTime1)
d.rt1$LogRT1 = log(d.rt1$ReactionTime1)


d.rt2$Connective = as.factor(d.rt2$Connective)
d.rt2$ID = as.factor(d.rt2$ID)
d.rt2$TrialID = as.factor(d.rt2$TrialID)
d.rt2$ReactionTime2 = as.numeric(d.rt2$ReactionTime2)
d.rt2$LogRT2 = log(d.rt2$ReactionTime2)
d.rt2$Truthvalue = as.factor(d.rt2$Truthvalue)
d.rt2$Sentenceform = as.factor(d.rt2$Sentenceform)

```


# simple model logRT1 ~ Connective
- not sognificant
```{r}

m.rt1 = lmer(ReactionTime1 ~ Connective + (1+Connective|ID) + (1|TrialID), data=d.rt1)
summary(m.rt1)

```


# simple model logRT1 ~ Connective
- not sognificant
```{r}

m.rt2 = lmer(LogRT2 ~ Connective*Truthvalue*Sentenceform + (1+Connective+Truthvalue+Sentenceform|ID) + (1|TrialID), data=d.rt2)
# saveRDS(m.ms.ma.full, "../models/model-task-para.rds")

# m <- readRDS("../models/m.rds")
summary(m.rt2)

```


# Second looking at double-processed data (no overal trial RT outlier removal)

```{r read in second data, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
d.rt1.double <- read.csv("../data/rt1_double_processed.csv")
d.rt2.double <- read.csv("../data/rt2_double_processed.csv")
```


```{r}
# Make everything into a factor
d.rt1.double$Connective = as.factor(d.rt1.double$Connective)
d.rt1.double$ID = as.factor(d.rt1.double$ID)
d.rt1.double$TrialID = as.factor(d.rt1.double$TrialID)
d.rt1.double$ReactionTime1 = as.numeric(d.rt1.double$ReactionTime1)
d.rt1.double$LogRT1 = log(d.rt1.double$ReactionTime1)


d.rt2.double$Connective = as.factor(d.rt2.double$Connective)
d.rt2.double$ID = as.factor(d.rt2.double$ID)
d.rt2.double$TrialID = as.factor(d.rt2.double$TrialID)
d.rt2.double$ReactionTime2 = as.numeric(d.rt2.double$ReactionTime2)
d.rt2.double$LogRT2 = log(d.rt2.double$ReactionTime2)
d.rt2.double$Truthvalue = as.factor(d.rt2.double$Truthvalue)
d.rt2.double$Sentenceform = as.factor(d.rt2.double$Sentenceform)

```


# simple model logRT1 ~ Connective
- not sognificant
```{r}

m.rt1.double = lmer(LogRT1 ~ Connective + (1+Connective|ID) + (1|TrialID), data=d.rt1.double)
# saveRDS(m.ms.ma.full, "../models/model-task-para.rds")

# m <- readRDS("../models/m.rds")
summary(m.rt1.double)

```


# simple model logRT1 ~ Connective
- ME only from 'and' to 'so'
- no 2-way with connective
- one three-way with connective
```{r}
m.rt2.double = lmer(LogRT2 ~ Connective*Truthvalue*Sentenceform + (1+Connective+Truthvalue+Sentenceform|ID) + (1|TrialID), data=d.rt2.double)
# saveRDS(m.ms.ma.full, "../models/model-task-para.rds")

# m <- readRDS("../models/m.rds")
summary(m.rt2.double)

```
